---
title:  Xanomeline Model Diagnostic Plots
author: Tim Bergsma
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  powerpoint_presentation:
    toc: FALSE
    keep_md: TRUE
    df_print: kable
    slide_level: 2
    reference_doc: template.pptx
editor_options: 
  chunk_output_type: console
---

```{r  setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(error = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(dev = c('png','pdf'))
knitr::opts_chunk$set(dpi = 600)
library(knitr)
library(magrittr)
library(dplyr)
library(tidyr)
library(wrangle)
library(ggplot2)
library(yamlet)
library(metaplot)
library(plotscale)
```

```{r data}
x <- io_csv('xanomeline-mod.csv', gz = TRUE)

x %<>% filter(cp > 1)
x %<>% ggready(parse = F)

```

```{r iso }
# iso() intends to enforce isometric aesthetics on a ggplot object.

isometric <- function()structure(list(), class = 'ggplot_isometric')

ggplot_add.ggplot_isometric <- function(object, plot, object_name){
  stopifnot('x' %in% names(plot$labels))
  stopifnot('y' %in% names(plot$labels))
  xrange <- range(na.rm = TRUE, plot$data[,plot$labels$x])
  yrange <- range(na.rm = TRUE, plot$data[,plot$labels$y])
  plot <- plot + expand_limits(x = yrange)
  plot <- plot + expand_limits(y = xrange)
  plot <- plot + theme(aspect.ratio = 1)
  plot
}
# p <- x %>% ggplot(aes(IPRED, DV)) + geom_point()
# p + isometric()

```

```{r symmetric }
# iso() intends to enforce isometric aesthetics on a ggplot object.

symmetric <- function()structure(list(), class = 'ggplot_symmetric')

ggplot_add.ggplot_symmetric <- function(object, plot, object_name){
  nms <- names(plot$labels)
  stopifnot('y' %in% nms)
  yrange <- range(na.rm = TRUE, plot$data[,plot$labels$y])
  plot <- plot + expand_limits(y = -yrange)
  plot
}
# p <- x %>% ggplot(aes(IWRES, TIME)) + geom_point()
# p + symmetric()

```

```{r functions}
isoplot <- function(
    x, 
    endpoint, versus, 
    alpha = 0.5, 
    log = T, 
    trans = 'identity',
    ... 
  ){
  
  x %>%
    ggplot(aes(x = {{ versus }}, y = {{ endpoint }} )) +
    geom_abline(slope = 1, intercept = 0) +
    geom_point(alpha = alpha) + 
    theme_bw() +
    theme(aspect.ratio = 1) + 
    geom_smooth(method = 'lm', formula = 'y ~ x') +
    isometric()
}


isopair <- function(
    x, 
    endpoint1, 
    versus1, 
    endpoint2, 
    versus2, 
    alpha = 0.5, # can be length 2
    trans = 'identity',
    ...
  ){
  alpha = rep(alpha, 2)
  trans = rep(trans, 2)
  multiplot(
    isoplot(x, {{ endpoint1 }}, {{versus1}}, alpha = alpha[[1]], trans = trans[[1]], ...),
    isoplot(x, {{ endpoint2 }}, {{versus2}}, alpha = alpha[[2]], trans = trans[[2]], ...) 
  )
}


trendplot <- function(
    x, 
    endpoint, versus, 
    alpha = 0.5, 
    log = T, 
    trans = 'identity',
    ref = 0,
    aspect = 1,
    symmetric = FALSE,
    ... 
){
  trans <- rep(trans, 2)
  p <- x %>%
    ggplot(aes(x = {{ versus }}, y = {{ endpoint }} )) +
    geom_hline(yintercept = ref) +
    geom_point(alpha = alpha) + 
    theme_bw() +
    theme(aspect.ratio = aspect) + 
    geom_smooth(method = 'lm', formula = 'y ~ x') +
    scale_y_continuous(trans = trans[[1]]) +
    scale_x_continuous(trans = trans[[2]])
  if(symmetric) p <- p + symmetric()
  p
}

trendpair <- function(
    x, 
    endpoint1, 
    versus1, 
    endpoint2, 
    versus2, 
    alpha = 0.5, # can be length 2
    trans1 = 'identity',
    trans2 = 'identity',
    ref = 0,
    aspect = 1,
    ...
){
  alpha = rep(alpha, 2)
  trans1 = rep(trans1, 2)
  trans2 = rep(trans2, 2)
  ref = rep(ref, 2)
  aspect = rep(aspect, 2)
  multiplot(
    trendplot(x, {{ endpoint1 }}, {{versus1}}, alpha = alpha[[1]], trans = trans1, aspect = aspect[[1]], ref= ref[[1]], ...),
    trendplot(x, {{ endpoint2 }}, {{versus2}}, alpha = alpha[[2]], trans = trans2, aspect = aspect[[2]], ref= ref[[2]], ...) 
  )
}

```

# Data Diagnostics

## Observations vs Individual Predictions by Arm, per Visit

```{r  DV-IPRED-ACTARM-VISIT, fig.width = 9.69, fig.height = 4.3, fig.cap = 'DV-IPRED-ACTARM-VISIT.png'}
#(
  x %>%
  ggplot(
    aes(IPRED, DV, color = ACTARM, pch = ACTARM)
  ) + 
    geom_point(alpha = .5, size = 1.5) +
   # geom_point(alpha = .2, size = 2) +
    scale_shape_manual(values = c(1,3))+
   # scale_shape_manual(values = c(0,4))+
    facet_wrap(~VISIT, ncol = 3) +
  theme_bw() +
  theme(
    aspect.ratio = 1, 
    legend.position = 'top', 
    legend.title = element_blank()
  ) +
  geom_abline(aes(slope = 1, intercept = 0)) +
  isometric()
  #) %>% devsize(3,3,verbose = T)
```

## Observations vs Individual Predictions

```{r  DV-PRED, fig.width = 3.65, fig.height = 3.52, fig.cap = "x %>% isoplot(DV, PRED)"}
x %>% isoplot(DV, PRED)
```

## Observations vs Individual Predictions, With/without Grids

```{r  DV-PRED-grids, fig.width = 7.3, fig.height = 3.52, fig.cap = "x %>% isoplot(DV, PRED)"}
multiplot(
x %>% isoplot(DV, PRED),
x %>% isoplot(DV, PRED) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
)
```

## Observations vs Population Predictions, Log-log

```{r  DV-PRED-LOG, fig.width = 3.65, fig.height = 3.52, fig.cap = "x %>% isoplot(DV, PRED, trans = 'log10')"}
x %>% isoplot(DV, PRED, trans = 'log10')
```

## Observations vs Population and Individual Predictions

```{r  DV-PRED-DV-IPRED, fig.width = 7.3, fig.height = 3.52, fig.cap = "x %>% isopair(DV, PRED, DV, IPRED)"}
x %>% isopair(DV, PRED, DV, IPRED)
```

## Observations vs Individual Predictions, Untransformed and Log-log

```{r  DV-IPRED-DV-IPRED-notrans-log, fig.width = 7.3, fig.height = 3.52, fig.cap = "x %>% isopair(DV, IPRED, DV, IPRED, trans = c('identity','log10'))"}
x %>% isopair(DV, IPRED, DV, IPRED, trans = c('identity','log10'))
```

## Individual Weighted Residuals vs Time

```{r  IWRES-TIME, fig.width = 3.65, fig.height = 3.52, fig.cap = "x %>% trendplot(IWRES, TIME) + symmetric()"}
x %>% trendplot(IWRES, TIME) + symmetric() #%>% devsize(3,3,verbose = T)
```

## Individual Weighted Residuals vs Time, Untransformed and Log-log

```{r  IWRES-TIME-IWRES-TIME-untrans-log, fig.width = 7.3, fig.height = 3.52, fig.cap = "x %>% trendpair(IWRES, TIME, IWRES, TIME, trans2 = c('identity', 'log10'), symmetric = TRUE)"}
x %>% trendpair(IWRES, TIME, IWRES, TIME, trans2 = c('identity', 'log10'), symmetric = TRUE)
```

## Individual Weighted Residuals vs Time, Full Grid vs Horizontal Grid

```{r  IWRES-TIME-IWRES-TIME-horizonal-grid, fig.width = 7.3, fig.height = 3.52 }
multiplot(
x %>% trendplot(IWRES, TIME)+ 
  scale_y_continuous(
    breaks = function(x)seq(
      from = round(x[[1]]), 
      to = round(x[[2]])
    )
  ) + 
  symmetric(),
x %>% trendplot(IWRES, TIME) + 
  theme(
    panel.grid.major.x = element_blank(), 
    panel.grid.minor.x = element_blank()
  )+ 
  scale_y_continuous(
    breaks = function(x)seq(
      from = round(x[[1]]), 
      to = round(x[[2]])
    )
  ) + symmetric()
)
```

## Individual Weighted Residuals vs Time, Various Horizontal Reference Lines

```{r  IWRES-TIME-IWRES-TIME-reference-lines, fig.width = 7.3, fig.height = 3.52}
multiplot(
x %>% 
  trendplot(IWRES, TIME) + 
  geom_hline(yintercept = c(2,4,-2,-4), linetype = 'dashed') + 
  scale_y_continuous(
    breaks = function(x)seq(
      from = round(x[[1]]), 
      to = round(x[[2]])
    )
  ) + symmetric()
,
x %>% trendplot(IWRES, TIME) + 
  geom_hline(yintercept = c(2,5,-2,-5), linetype = 'dashed') + 
    scale_y_continuous(
    breaks = function(x)seq(
      from = round(x[[1]]), 
      to = round(x[[2]])
    )
  ) + symmetric()
)
```

## Individual Plots

```{r  individuals, fig.width = 8.02, fig.height = 6.25}
(
  x %>% 
  filter(desolve(ID) <= 1068) %>%
  mutate(Profile = as.factor(dosenum)) %>%
  ggplot(aes(x = TAD, group = dosenum)) +
  geom_point(
    aes(
      y = DV, 
      shape = Profile
    ),    
    alpha = 1, 
    color = "#144A90") + # blue
  geom_line(
    aes(y = PRED),  
    alpha = 1, 
    color = "#AA00FF" # violet
  ) + 
  geom_line(
    aes(y = IPRED), 
    alpha = 1, 
    color = "#4CB54F" # green
  ) + 
  theme_bw() +
  facet_wrap(~desolve(ID), ncol = 7) +
  theme(legend.position  = 'top')
)  #%>% devsize(1,1, verbose = T)

```

## Individual Plots -- Detailed Panel Strip

```{r  individuals-detail, fig.width = 8.02, fig.height = 5.05}
(
  x %>% 
  filter(desolve(ID) <= 1029) %>%
  mutate(Profile = as.factor(dosenum)) %>%
  ggplot(aes(x = TAD, group = dosenum)) +
  geom_point(
    aes(
      y = DV, 
      shape = Profile
    ),    
    alpha = 1, 
    color = "#144A90") + # blue
  geom_line(
    aes(y = PRED),  
    alpha = 1, 
    color = "#AA00FF" # violet
  ) + 
  geom_line(
    aes(y = IPRED), 
    alpha = 1, 
    color = "#4CB54F" # green
  ) + 
  theme_bw() +
  facet_wrap(
    . ~'Study 01' + 
      ID + 
      gsub(
        'Xanomeline','',
        as.character(ACTARM)
      ) + 
      'Transdermal'
    , 
    labeller = label_value, 
    ncol = 7
  ) +
  theme(
    legend.position  = 'top', 
    strip.background = element_rect(linetype = 'blank')
  )
) # %>% devsize(1,1, verbose = T)

```





