---
title: "Scripted PDF"
author: "Tim Bergsma"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    extra_dependencies:
      booktabs,
      longtable,
      array,
      multirow,
      wrapfig,
      float,
      colortbl,
      pdflscape,
      tabu,
      threeparttable,
      threeparttablex,
      makecell,
      xcolor
vignette: >
  %\VignetteIndexEntry{Scripted PDF}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

The point of this exercise is to demonstrate flexible rendering of 
subscripts and superscripts.  We want to write expressions for 
column labels an units that are fairly readable as they are, and
yet can be easily rendered with equivalent results in 
plotmath, html, or pdf.

First we load some packages.

```{r}
library(magrittr)
library(ggplot2)
library(tablet)
library(yamlet)
```

We create some toy data.

```{r}
x <- data.frame(time = 1:10, work = (1:10)^1.5)
x %<>% decorate('
 time: [ Time_cum.^alpha, h ]
 work: [ Work_total_obs, kg*m^2/s^2 ]
')
x %>% decorations
```

Notice that the label for "work" has nested subscripts:  "obs" is
subordinate to "total", and the result is subordinate to "Work".

For "time", the word "Time" is followed by a subscript "cum".
Then, "cum" is followed by the superscript "alpha".
In this case "alpha" to be the superscript of "Time_cum" not just "cum".
We include the dot to explicitly terminate the "cum"
subscript, which causes "alpha" to be the superscript of
the combined result.

How does this look when we plot it?
```{r}
x %>% resolve %>% ggplot(aes(time, work)) + geom_point()
```

We get verbatim labels and units as substitutes for column names.

Next, we use scripted() instead of resolve() to indicate 
that the labels should be understood as
potentially having subscripts and superscripts.
For this to work well, units should be constructed
using *, /, and ^ (even though the "units"
package supports other encodings).

```{r}
x %>% scripted %>% ggplot(aes(time, work)) + geom_point()
```

In the background, scripted() is writing "expression" attributes
(consumed by ggplot() ) and "title" attributes (consumed by tablet() ).
We illustrate the latter.

```{r}
x %>% scripted %>% tablet %>% as_kable
```

In summary, we have decorated our data with labels and 
units containing markup for subscripts and superscripts.
If everything goes well, these render similarly 
in figures and tables.  They also render similarly in
html and pdf.  Please see the html version of this document.

